The Early Eocene Climate Optimum reconstructed with Bayesian multi-proxy integration

### 2022-08-11 manuscript/manuscript.Rmd
Cleaning repo
manuscript notes, option to fix sdy

The early Eocene climate optimum (EECO, ~ 53 - 49 million years ago) is thought to have been the warmest episode ever since. The EECO has received much attention due to its use as an analogue for extreme warming scenarios, depicting a possible future high-CO2 world. To reconstruct the climate of EECO, a variety of proxy data and Earth System Models have been used. Early modelers have soon noticed that their models struggle to reproduce the climate indicated by proxy data: Without unrealistically high concentrations of atmospheric CO2, early climate models struggled to reproduce the equitable polar climates indicated by early Eocene proxy data, e.g. the occurrence of mangroves in polar latitudes, and d18O data from x indicates a temperature of ~ y at latitude z.

In order to properly compare model output to proxy data, latitudinal temperature patterns need to be reconstructed from noisy data (impossible to properly compare model gradients with scattered proxy points "by eye")

>>>>

The early Eocene climate optimum (EECO, ~ 53 - 49 million years ago) is thought to have been the warmest episode of [...]. The EECO is of particular interest due to its use as an analogue for extreme warming scenarios, depicting a possible future high-CO2 world. To reconstruct the climate of EECO, a variety of proxy data and Earth System Models have been used. Early modelers have soon noticed that their models struggle to reproduce the climate indicated by proxy data: Without unrealistically high concentrations of atmospheric CO2, early climate models were unable to reproduce the equitable polar climates indicated by early Eocene proxy data, e.g. the occurrence of mangroves in polar latitudes, and d18O data from x indicates a temperature of ~ y at latitude z.

Contrast between complex, process-driven Earth system models and very simplistic use of the proxy data as local temperature estimates or global averages. Here, we introduce a proxy-driven modelling approach of relatively low complexity, which places the available chemical and biological proxy data in a single climate model.
In order to properly compare model output to proxy data, latitudinal temperature patterns need to be reconstructed from noisy data (impossible to properly compare model gradients with scattered proxy points "by eye")
